Investigating motor vehicle collisions in Toronto: Installing more red light cameras do not stop traffic accidents from happening

### 2022-02-06 outputs/paper/paper.Rmd
fix3

The Motor Vehicle Collisions data of Toronto provides a record of the past severe vehicle collisions in Toronto which result in injuries or deaths, and with the help of statistical data analysis, it can possibly offer insights on the ways to reduce this type of accident. This report investigates the motor vehicle collisions in Toronto from 2006 to 2020, with a focus on the ones result in injuries or deaths. In Section \@ref(data), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of traffic accidents with respect to their fatality and their location. Then, I discussed about the possible implications of these visualizations. By Figure 2, I concluded that motor vehicle collisions are always more likely to occur at intersections. Thus, 
Continuing

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data
## Motor Vehicle Collisions involving Killed or Seriously Injured Persons
To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal [@citedata]. Provided by the Toronto Police Service, this dataset includes all vehicle collisions occurred in Toronto between 2006 and 2020 that result in injuries or deaths. One bias aroused by this dataset is the offset from the real accident location. For ethical reason and the protection of the parties involved in the accidents that are considered to be criminal cases, the location of those criminal accidents are offset to the closest intersection node [@citecollisiondata]. As a result, the location of those accidents specific to neighborhoods might not be accurate. Therefore, I will use the variable district instead of neighbourhood to categorize the accidents later in the report to avoid this bias.

This dataset contains 16840 observations of collisions and 54 variables with all kinds of information of each collision. This report focuses on four of these variables which are: year, accident location, district, and fatality. By using R [@citeR], and R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr], an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load motor vehicle collision data and clean
vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
  clean_names() %>%
  drop_na(year) %>%
  filter(acclass != "Property Damage Only")
```

@@

## Red Light Cameras
The number of red light camera instilled in Toronto from 2007 to 2020 is obtained from the dataset "Red Light Cameras" found in the Toronto Open Data Portal [@citedata]. This dataset is shared by the Transportation Services and it covers the locations of red light cameras in Toronto. In its introduction, this dataset claims that red light cameras can improve the safety of the drivers and pedestrians by monitoring the behavior of the drivers [@citecameradata]. It contains 199 observations and 32 variables. This dataset's downside is that it is missing values in lots of its variables. To reduce errors generated from this downside, I will only utilize the variables with no missing values. R [@citeR], R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr] are used to produce an extract of the cleaned dataset (Table \@ref(tab:dadaextract1)).

```{r dadaextract1, message=FALSE, warning=FALSE, echo=FALSE}
# Load red light camera data and clean
red_light_camera <- read_csv(here::here("inputs/data/red_light_camera.csv")) %>%
  clean_names() %>%
  select(activation_date, district) %>%
  drop_na(activation_date) %>%
  drop_na(district)

@@

# Discussion
The analysis of the two datasets show no relationship between the cases of motor vehicle collisions in Toronto and the number of red light cameras. From the comparision between Figure 3 and Figure 4, we see that in none of the districts did the number of vehicle collisions followed the trend of the installations of the red light cameras. For stance, between 2016 and 2017, all four districts made the installations of lots of new red light cameras, but none of the data in Figure 4 seemed to reflect that action. The implication of Section \@ref(data) is that: installing more red cameras does not change the number of traffic accidents in Toronto, nor does it influence the fatality of the accidents. A very likely reason to explain this situation is that red light cameras only concerned the drivers who already followed the rule most of the time, but not the rule-breaking drivers. Another possible reason would be that at intersections with red light cameras, more drivers would intend to stop rather than to proceed when they see the light is about to turn red, resulting in a higher chance of rear end accidents [@cite2].

In recent years, complaints have been raised about whether red light cameras are used mainly to prevent injuries and deaths from traffic accidents or earn revenues for the government [@cite3]. Measures should be taken to improve the effectiveness of this system and to ensure the safety of drivers and pedestrians who are the victims of the vehicle accidents. For instance, drivers should be more notified on the existence of red light cameras, instead of knowing them after they are fined, because red light cameras is supposed to prevent accidents but not to punish people after they cause accidents. Government could also seek for other solutions to reduce these accidents. Toronto and East York has done a great job reducing the number of motor vehicle collisions, which means other district could adopt similar traffic controls and see how it works in a few years.

>>>>

The Motor Vehicle Collisions data of Toronto provides a record of the past severe vehicle collisions in Toronto which result in injuries or deaths, and with the help of statistical data analysis, it can possibly offer insights on the ways to reduce this type of accident. This report investigates the motor vehicle collisions in Toronto from 2006 to 2020, with a focus on the ones result in injuries or deaths. In Section \@ref(motor-vehicle-collisions-involving-killed-or-seriously-injured-persons), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of traffic accidents with respect to their fatality and their location. Then, I discussed about the possible implications of these visualizations. Figure 1 showed that there was a descending trend of traffic accidents from 2006 to 2020. By Figure 2, I concluded that motor vehicle collisions are always more likely to occur at intersections. Thus, I brought out the Red Light Cameras dataset in Section \@ref(red-light-cameras) to see whether if the number of red light cameras has an effect on the number of traffic accidents with serious injuries and deaths. The result from the datasets indicates that there is no clear relationship between this two factors, suggesting that just by installing more red light cameras cannot prevent traffic accidents from happening at intersections. More implications and further discussions on this result is presented in Section \@ref(discussion).

# Data
## Motor Vehicle Collisions involving Killed or Seriously Injured Persons
To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal [@citedata]. Provided by the Toronto Police Service, this dataset includes all vehicle collisions occurred in Toronto between 2006 and 2020 that result in injuries or deaths. One bias aroused by this dataset is the offset from the real accident location. For ethical reason and the protection of the parties involved in the accidents that are considered to be criminal cases, the location of those criminal accidents are offset to the closest intersection node [@citecollisiondata]. As a result, the location of those accidents specific to neighborhoods might not be accurate. Therefore, I will use the variable district instead of the variable neighbourhood to categorize the accidents later in the report to avoid the bias.

This dataset contains 16840 observations of collisions and 54 variables with all kinds of information of each collision. This report focuses on four of these variables which are: year, accident location, district, and fatality. By using R [@citeR], and R packages "tidyverse" [@citetidyverse], "dplyr" [@citedplyr], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr], an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load motor vehicle collision data and clean
vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
  clean_names() %>%
  drop_na(year) %>%
  filter(acclass != "Property Damage Only")
```

@@

## Red Light Cameras
The number of red light camera instilled in Toronto from 2007 to 2020 is obtained from the dataset "Red Light Cameras" found in the Toronto Open Data Portal [@citedata]. This dataset is shared by the Transportation Services and it covers the locations of red light cameras in Toronto. In its introduction, this dataset claims that red light cameras can improve the safety of the drivers and pedestrians by monitoring the behavior of the drivers [@citecameradata]. It contains 199 observations and 32 variables. This dataset's downside is that it is missing values in lots of its variables. To reduce errors generated from this downside, I will only utilize the variables with no missing values. R [@citeR], R packages "tidyverse" [@citetidyverse], "dplyr" [@citedplyr], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr] are used to produce an extract of the cleaned dataset (Table \@ref(tab:dadaextract1)).

```{r dadaextract1, echo=FALSE, message=FALSE, warning=FALSE}
# Load red light camera data and clean
red_light_camera <- read_csv(here::here("inputs/data/red_light_camera.csv")) %>%
  clean_names() %>%
  select(activation_date, district) %>%
  drop_na(activation_date) %>%
  drop_na(district)

@@

# Discussion
The analysis of the two datasets show no relationship between the cases of motor vehicle collisions in Toronto and the number of red light cameras. From the comparison between Figure 3 and Figure 4, we see that in none of the districts did the number of vehicle collisions followed the trend of the installations of the red light cameras. For stance, between 2016 and 2017, all four districts made the installations of lots of new red light cameras, but none of the data in Figure 4 seemed to reflect that action. The implication of Section \@ref(data) is that: installing more red cameras does not change the number of traffic accidents in Toronto, nor does it influence the fatality of the accidents. A very likely reason to explain this situation is that red light cameras only concerned the drivers who already followed the rule most of the time, but not the rule-breaking drivers. Another possible reason would be that at intersections with red light cameras, more drivers would intend to stop rather than to proceed when they see the light is about to turn red, resulting in a higher chance of rear end accidents [@cite2].

In recent years, complaints have been raised about whether red light cameras are used mainly to prevent injuries and deaths from traffic accidents or earn revenues for the government [@cite3]. Measures should be taken to improve the effectiveness of this system and to ensure the safety of drivers and pedestrians who are the victims of the vehicle accidents. For instance, drivers should be more notified on the existence of red light cameras, instead of knowing them after they are fined, because red light cameras is supposed to prevent accidents but not to punish people after they cause accidents. Government could also seek for other solutions to reduce these accidents. Toronto and East York has done a great job reducing the number of motor vehicle collisions, which means that other district could adopt similar traffic controls and see how it works in a few years.
